### 2014-04-06 cs250-hw/hw1.tex
c1

\documentclass{article}
\usepackage{geometry}
\usepackage[namelimits,sumlimits]{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage[cm]{fullpage}
\newcommand{\tab}{\hspace*{5em}}
\newcommand{\conj}{\overline}
\newcommand{\dd}{\partial}
\newcommand{\ep}{\epsilon}
\newcommand{\openm}{\begin{pmatrix}}
\newcommand{\closem}{\end{pmatrix}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Null}{null}
\newcommand{\nc}{\newcommand}
\newcommand{\rn}{\mathbb{R}}
\newcommand{\zn}{\mathbb{Z}}
\nc{\cn}{\mathbb{C}}
\nc{\ssn}[1]{\subsubsection*{#1}}
\nc{\inner}[2]{\langle #1,#2\rangle}
\nc{\h}[1]{\widehat{#1}}
\nc{\tl}[1]{\widetilde{#1}}
\nc{\norm}[1]{\left\|{#1}\right\|}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\begin{document}
Name: Hall Liu

Date: \today 
\vspace{1.5cm}

\subsection*{1}

\subsection*{3}
\ssn{b}
I have a general policy against running unknown scripts on my computer, and since I know that I have a large part of the Python toolchain set up already, I decided to read the setup script (\verb|toolkit.sh|) and replicate the necessary parts.

The first difficulty came when trying to install the packages. I ran into a syntax error while installing czipfile via pip, and this turned out to be because my system defaults to Python 3. Deleting the virtualenv and re-running mkvirtualenv with the option \verb|--python=/usr/bin/python2.7| fixed the issue.

The Python version problem came up again when creating the bin/notebook file. I modified the line to start with \verb|python2.7 /usr/bin/ipython| to force the script to use 
\end{document}