### 2014-04-11 stat246/hw2.tex
s2

\documentclass{article}
\usepackage{geometry}
\usepackage[namelimits,sumlimits]{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage[cm]{fullpage}
\newcommand{\tab}{\hspace*{5em}}
\newcommand{\conj}{\overline}
\newcommand{\dd}{\partial}
\newcommand{\ep}{\epsilon}
\newcommand{\openm}{\begin{pmatrix}}
\newcommand{\closem}{\end{pmatrix}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Null}{null}
\newcommand{\nc}{\newcommand}
\newcommand{\rn}{\mathbb{R}}
\nc{\cn}{\mathbb{C}}
\nc{\ssn}[1]{\subsubsection*{#1}}
\nc{\inner}[2]{\langle #1,#2\rangle}
\nc{\h}[1]{\widehat{#1}}
\nc{\tl}[1]{\widetilde{#1}}
\nc{\norm}[1]{\left\|{#1}\right\|}
\begin{document}

Name: Hall Liu

Date: \today 
\vspace{1.5cm}
\subsection*{1}
The mgf of $Y$ is $M_Y(t)=E(e^{t^TY})=E(e^{t^TAx})=E(e^{(A^Tt)^Tx})=M_X(A^Tt)$, which we can write as $\exp\left(t^TA\mu+\frac{1}{2}t^TA\Sigma A^Tt\right)$. This is the mgf of a normal RV with mean $A\mu$ and covariance $A\Sigma A^T$, so $Y$ must be also distributed as such.

\subsection*{2}
\ssn{a}
We have that $f(x,y)=f_Y(y|x)f_X(x)$, so $f(x,y)$ is jointly normal iff this product has the form of one. In (b), we show that there exists a particular definition for $Z$ such that $f_Z(z|x)=f_Y(y|x)$ and $X$ and $Z$ are jointly normal. Thus, since $f(x,z)$ and $f(x,y)$ are equivalent, $f(x,y)$ is normal.
\ssn{b}
Let $Z=AX+b+U$, $U\sim N(0,Q)$ and independent of $X$. Then, if given a value for $X$, we have $Z=Ax+b+U$, so consequently we have $f(z|x)=N(Ax+b, Q)$, the same as that of $f(y|x)$.

The joint mgf of $X$ and $Z$ can be written as 
\[E\left(\exp\left(\openm t_1^T&t_2^T\closem\openm X\\Z\closem\right)\right)=E(\exp(t_1^TX+t_2^T(AX+b+U)))=E(\exp((t_1+A^Tt_2)^TX)\exp(t_2^Tb)\exp(t_2^TU))\]
Resolving the expected value and expressing the result using the mgfs of $X$ and $U$, we have
\[\exp\left((t_1+A^Tt_2)^T\mu+t_2^Tb+\frac{1}{2}(t_1+A^Tt_2)^T\Sigma (t_1+A^Tt_2)+\frac{1}{2}t_2^TQt_2\right)\]
Now, we note that 
\[(t_1+A^Tt_2)^T\mu+t_2^Tb=\openm t_1^T&t_2^T\closem\openm\mu\\A\mu+b\closem\text{ and}\]
\[(t_1+A^Tt_2)^T\Sigma (t_1+A^Tt_2)+t_2^TQt_2=\openm t_1^T&t_2^T\closem\openm \Sigma &\Sigma A^T\\A\Sigma &A\Sigma A^T+Q\closem\openm t_1\\t_2\closem\]
Thus, if we let $t=\openm t_1\\t_2\closem$, $\mu'=\openm\mu\\A\mu+b\closem$, and $\Sigma'=\openm \Sigma &\Sigma A^T\\A\Sigma &A\Sigma A^T+Q\closem$, we have that the joint mgf of $X,Z$ is $\exp\left(t^T\mu'+\frac{1}{2}t^T\Sigma't\right)$, which means that it's multivariate normal with mean $\mu'$ and covariance $\Sigma'$.
\ssn{c}
The marginal is simply restricting the joint distribution to the relevant components. In the case of $Y$, its marginal distribution is then $N(A\mu+b, A\Sigma A^T+Q)$. For the density of $X$ conditional on $Y=y$, (2.81) and (2.82) in the book give us that it's normal with $\mu_{x|y}=\mu+\Sigma A^T(A\Sigma A^T+Q)^{-1}(y-A\mu+b)$ and $\Sigma_{x|y}=\Sigma-\Sigma A^T(A\Sigma A^T+Q)^{-1}A\Sigma$
\end{document}