### 2014-04-13 stat246/hw2.tex
s2 done

\documentclass{article}
\usepackage{geometry}
\usepackage[namelimits,sumlimits]{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage[cm]{fullpage}
\newcommand{\tab}{\hspace*{5em}}
\newcommand{\conj}{\overline}
\newcommand{\dd}{\partial}
\newcommand{\ep}{\epsilon}
\newcommand{\openm}{\begin{pmatrix}}
\newcommand{\closem}{\end{pmatrix}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Null}{null}
\DeclareMathOperator{\flatv}{flat}
\newcommand{\nc}{\newcommand}
\newcommand{\rn}{\mathbb{R}}
\nc{\cn}{\mathbb{C}}
\nc{\ssn}[1]{\subsubsection*{#1}}
\nc{\inner}[2]{\langle #1,#2\rangle}
\nc{\h}[1]{\widehat{#1}}
\nc{\tl}[1]{\widetilde{#1}}
\nc{\norm}[1]{\left\|{#1}\right\|}
\begin{document}

Name: Hall Liu

Date: \today 
\vspace{1.5cm}
\subsection*{1}
The mgf of $Y$ is $M_Y(t)=E(e^{t^TY})=E(e^{t^TAX})=E(e^{(A^Tt)^TX})=M_X(A^Tt)$, which we can write as $\exp\left(t^TA\mu+\frac{1}{2}t^TA\Sigma A^Tt\right)$. This is the mgf of a normal RV with mean $A\mu$ and covariance $A\Sigma A^T$, so $Y$ must be also distributed as such.

\subsection*{2}
\ssn{a}
We have that $f(x,y)=f_Y(y|x)f_X(x)$, so $f(x,y)$ is jointly normal iff this product has the form of one. In (b), we show that there exists a particular definition for $Z$ such that $f_Z(z|x)=f_Y(y|x)$ and $X$ and $Z$ are jointly normal. Thus, since $f(x,z)$ and $f(x,y)$ are equivalent, $f(x,y)$ is normal.
\ssn{b}
Let $Z=AX+b+U$, $U\sim N(0,Q)$ and independent of $X$. Then, if given a value for $X$, we have $Z=Ax+b+U$, so consequently we have $f(z|x)=N(Ax+b, Q)$, the same as that of $f(y|x)$.

The joint mgf of $X$ and $Z$ can be written as 
\[E\left(\exp\left(\openm t_1^T&t_2^T\closem\openm X\\Z\closem\right)\right)=E(\exp(t_1^TX+t_2^T(AX+b+U)))=E(\exp((t_1+A^Tt_2)^TX)\exp(t_2^Tb)\exp(t_2^TU))\]
Resolving the expected value and expressing the result using the mgfs of $X$ and $U$, we have
\[\exp\left((t_1+A^Tt_2)^T\mu+t_2^Tb+\frac{1}{2}(t_1+A^Tt_2)^T\Sigma (t_1+A^Tt_2)+\frac{1}{2}t_2^TQt_2\right)\]
Now, we note that 
\[(t_1+A^Tt_2)^T\mu+t_2^Tb=\openm t_1^T&t_2^T\closem\openm\mu\\A\mu+b\closem\text{ and}\]
\[(t_1+A^Tt_2)^T\Sigma (t_1+A^Tt_2)+t_2^TQt_2=\openm t_1^T&t_2^T\closem\openm \Sigma &\Sigma A^T\\A\Sigma &A\Sigma A^T+Q\closem\openm t_1\\t_2\closem\]
Thus, if we let $t=\openm t_1\\t_2\closem$, $\mu'=\openm\mu\\A\mu+b\closem$, and $\Sigma'=\openm \Sigma &\Sigma A^T\\A\Sigma &A\Sigma A^T+Q\closem$, we have that the joint mgf of $X,Z$ is $\exp\left(t^T\mu'+\frac{1}{2}t^T\Sigma't\right)$, which means that it's multivariate normal with mean $\mu'$ and covariance $\Sigma'$.
\ssn{c}
The marginal is simply restricting the joint distribution to the relevant components. In the case of $Y$, its marginal distribution is then $N(A\mu+b, A\Sigma A^T+Q)$. For the density of $X$ conditional on $Y=y$, (2.81) and (2.82) in the book give us that it's normal with $\mu_{x|y}=\mu+\Sigma A^T(A\Sigma A^T+Q)^{-1}(y-A\mu+b)$ and $\Sigma_{x|y}=\Sigma-\Sigma A^T(A\Sigma A^T+Q)^{-1}A\Sigma$
\subsection*{3}
The likelihood of the data is 
\[\prod_{i=1}^N\frac{C}{\sqrt{|\Sigma|}}\exp\left(-\frac{1}{2}(X_i-\mu_{k_i})^T\Sigma^{-1}(X_i-\mu_{k_i})\right)\pi_{k_i}\]
where $C$ is a constant that depends on none of the parameters. We have immediately that $\h{\pi}_{k_i}$ is $\frac{\#i}{N}$, where $\#i$ is the number of occurrances of class $i$ in the data. Taking the loglikelihood (and ignoring the constant $C$ and the factor of $\frac{1}{2}$), we have
\[\sum_{i=1}^N-\log|\Sigma|+(X_i-\mu_{k_i})^T\Sigma^{-1}(X_i-\mu_{k_i})+\log(\pi_{k_i})\]
To minimize this expression over the $\mu_k$, write the expression with an inner sum over points with the same class:
\[-N\log|\Sigma|+\sum_{k=1}^K\sum_{X:Y(X)=k}(X-\mu_k)^T\Sigma^{-1}(X-\mu_k)\]
The $\pi_{k_i}$ terms have been discarded because they are not dependent on the parameters. Now, we can simply minimize over each class, and find that $\mu_k$ is the sample mean over all points that are in class $k$.

Now, let $S_k$ be the sample covariance over class $k$, or $\sum_{i=1}^{\#k}(X_i-\conj{X})(X_i-\conj{X})^T$. Then, we can write the log-likelihood (divided out by $N$) as
\[-\log|\Sigma|+\frac{1}{N}\sum_{i=1}^N\tr\left((X_i-\mu_{k_i})(X_i-\mu_{k_i})^T\Sigma^{-1}\right)=-\log|\Sigma|+\frac{1}{N}\sum_{k=1}^K\tr\left(\#kS_k\Sigma^{-1}\right)=-\log|\Sigma|+\tr\left(\left(\sum_{k=1}^K\frac{\#k}{N}S_k\right)\Sigma^{-1}\right)\]
It was shown that minimizing an expression of the form $-\log|\Sigma|+\tr(A\Sigma^{-1})$ over $\Sigma$ is given by $\Sigma=A$, so we have that the MLE for the pooled covariance is $\sum_{k=1}^K\frac{\#k}{N}S_k$
\subsection*{4}
Let $g(\eta)=\exp(A(\eta))$. Then, we have $f(x;\eta)=\frac{h(x)}{g(\eta)}e^{\eta^TT(x)}$, and integrating this over $\rn^n$ (the space $x$ lies in) gives $1$, so we have $g(\eta)=\int_{\rn^n}h(x)e^{\eta^TT(x)}dx$.

Taking the first derivative of $g$ gives $\nabla g(\eta)=e^{A(\eta)}\nabla A(\eta)$ and 
\[\nabla g(\eta)=\int h(x)e^{\eta^TT(x)}T(x)dx\implies\nabla A(\eta)=\int h(x)e^{\eta^TT(x)-A(\eta)}T(x)=E(T(x))\]
Using the notation $\nabla^2 g$ for the second derivative, we have that $\nabla^2g(\eta)=e^{A(\eta)}\left(\nabla^2A(\eta)+\nabla A(\eta)\nabla A(\eta)^T\right)$ and
\[\nabla^2 g(\eta)=\int h(x)e^{\eta^TT(x)}T(x)T(x)^Tdx\]
Combining the two expressions gives us 
\[e^{A(\eta)}\left(\nabla^2 A(\eta)+E(T(x))E(T(x)^T)\right)=\int h(x)e^{\eta^TT(x)}T(x)T(x)^Tdx\]
Dividing out by $e^{A(\eta)}$ and note that the RHS is equal to $E(T(x)T(x)^T)$, so subtracting off $E(T(x))E(T(x)^T)$ from both sides gives $\nabla^2A(\eta)=E(T(x)T(x)^T)-E(T(x))E(T(x)^T)$, which is the covariance of $T(x)$.
\subsection*{5}
The density of the multivariate normal distribution can be written as 
\[\frac{1}{(2\pi)^{d/2}|\Sigma|}\exp\left(-\frac{1}{2}\tr\left((x-\mu)(x-\mu)^T\Sigma^{-1}\right)\right)\]
We can let $(2\pi)^{-d/2}$ be $h(x)$ and put $|\Sigma^{-1}|$ into $g(\mu, \Sigma)$, so we are left with considering the exponential term. Expanding out the exponent, we get
\[\frac{-1}{2}\tr\left(xx^T-\mu x^T-x\mu^T+\mu\mu^T)\Sigma^{-1}\right)=-\frac{1}{2}\left(\tr(xx^T\Sigma^{-1})-2\tr(x\mu^T\Sigma^{-1})+\tr(\mu\mu^T\Sigma^{-1})\right)\]
Now, let $g(\mu,\Sigma)=|\Sigma^{-1}|\exp\left(-\frac{1}{2}\mu^T\Sigma^{-1}\mu\right)$, so we can now discard the last term from the above expression and try to find what $T(x)$ is. We're left with
\[-\frac{1}{2}\tr(xx^T\Sigma^{-1})+\mu^T\Sigma^{-1}x\]

Define $\flatv(A)$ for a $n\times m$ matrix $A$ to be a $nm$-long column vector that consists of the rows of $A$ laid out in sequential order. Then, we have that $\tr(AB)=\flatv(A)^T\flatv(B^T)$ for matrices $A,B$. Let $\eta=\openm\Sigma^{-1}\mu\\-\frac{1}{2}\flatv(\Sigma^{-1})\closem$ and let $T(x)$ be $\openm x\\\flatv(xx^T)\closem$. Then, we have that
\[\eta^TT(x)=\openm\mu^T\Sigma^{-1}&-\frac{1}{2}\flatv(\Sigma^{-1})^T\closem\openm x\\\flatv(xx^T)\closem=\mu^T\Sigma^{-1}x-\frac{1}{2}\tr(\Sigma^{-1}xx^T)\]
which corresponds to the terms inside the exponent that were not absorbed by $g(\mu,\Sigma)$. Thus, we have that the dimension of $T(x)$ and $\eta$ is $d+d^2$.

In summary, $h(x)=(2\pi)^{-d/2}$, $g(\mu,\Sigma)=|\Sigma^{-1}|\exp\left(-\frac{1}{2}\mu^T\Sigma^{-1}\mu\right)$, $\eta=\openm\Sigma^{-1}\mu\\-\frac{1}{2}\flatv(\Sigma^{-1})\closem$, and $T(x)=\openm x\\\flatv(xx^T)\closem$.
\subsection*{6}
\ssn{a}
If we write $x^{\alpha-1}=e^{(\alpha-1)\log(x)}$, the density of the gamma distribution becomes
\[\frac{1}{\Gamma(\alpha)\beta^\alpha}\exp\left(-\frac{x}{\beta}+(\alpha-1)\log(x)\right)=\frac{1}{\Gamma(\alpha)\beta^\alpha}\exp\left(\openm-\beta^{-1}&\alpha-1\closem\openm x\\\log(x)\closem\right)\]
Thus, we have $\eta=\openm-\beta^{-1}\\\alpha-1\closem$, $T(x)=\openm x\\\log(x)\closem$, $g(\eta)=\frac{1}{\Gamma(\alpha)\beta^\alpha}$, and $h(x)=1$.

By (2.228), the ML estimator for $\eta$ satisfies
\[-\nabla \log g(\eta)=\frac{1}{N}\sum_{i=1}^NT(x_i)\]
In terms of the components of $\eta$, we have $g(\eta)=\frac{(-\eta_1)^{\eta_2+1}}{\Gamma(\eta_2+1)}$. The negative log of this is $\log\Gamma(\eta_2+1)-(\eta_2+1)\log(-\eta_1)$, and its gradient is $\openm\psi^{(0)}(\alpha)-\log\beta\\-\frac{\alpha}{\beta}\closem$, where $\psi^{(0)}$ is the polygamma function.

Then, the pair of equations for $\alpha$ and $\beta$ are 
\begin{align*}
    \psi^{(0)}(\alpha)-\log\beta&=\conj{x}\text{ and}\\
    -\frac{\alpha}{\beta}&=\frac{1}{N}\sum_{i=1}^N\log x_i\\
\end{align*}
Due to the lack of a closed-form inverse for the polygamma function, there probably exists no closed-form solution to this system.
\ssn{b}
The notation got really confusing because $\beta$ is used as a scale parameter here rather than the rate parameter that it usually stands for. I'm just going to flip it around and let $\theta=\beta^{-1}$. If we hold $\alpha$ fixed, we can write the gamma-$\alpha$ distribution with parameter $\theta$ in the exponential family form as
\[\frac{x^{\alpha-1}}{\Gamma(\alpha)}\theta^\alpha e^{\theta(-x)}\]
with $g(\theta)=\theta^\alpha$ and $T(x)=-x$. Then, by (2.229), the conjugate prior for $\theta$ has the form
\[f(\chi, \nu)\theta^{\alpha\nu}e^{\nu\theta\chi}\]
where $\chi$ is a hyperparameter. If we set $\nu=\frac{a-1}{\alpha}$, this has the functional form of a gamma distribution with parameters $a,b$($b$ is a rate parameter), where $b=-\frac{\chi(a-1)}{\alpha}$. Thus, the conjugate prior of $\theta$ is a gamma distribution, which means that the conjugate prior of $\beta$ is an inverse gamma distribution.

If we work in terms of $\theta$, the likelihood for the data $X_i$ is 
\[\frac{\theta^{N\alpha}}{\Gamma(\alpha)^N}\prod_{i=1}^NX_i^{\alpha-1}e^{-X_i\theta}\]
Multiplying this by the conjugate prior produces
\[\left(\frac{\prod_{i=1}^NX_i^{\alpha-1}}{\Gamma(\alpha)^N}\cdot\frac{b^a}{\Gamma(a)}\right)\theta^{N\alpha+a-1}e^{-\theta\sum_{i=1}^NX_i-b\theta}\]
Up to some normalizing constant that absorbs all the stuff in the front in the parens, this is another gamma distribution with shape $a+N\alpha$ and rate $b+\sum X_i$, so the posterior mean of $\theta$ is $\frac{a+N\alpha}{b+\sum X_i}$, which means that the posterior mean of $\beta$ is the reciprocal of this.

\subsection*{7}
Since the number of samples from each distribution is equal, the decision can be made by just seeing which of the likelihoods is greater. Doing so, we obtain the following plot of the decision boundaries, each a locus of solutions to a quadratic $x^TAx+b^Tx+c=0$. The black curve is the boundary obtained from using the true parameters, the red curve the non-pooled covariance, and the blue line is the linear boundary from the pooled covariance (still a quadratic, but $A=0$).

\includegraphics[width=0.6\textwidth]{hw2_files/boundaries.png}

The red curve from the non-pooled covariance is pretty different from the true decision boundary -- it classifies a lot of points to the side as being from the first distribution instead of the second, as the true decision boundary does. However, the linear decision boundary looks a lot more reasonable -- if we had to draw a line separating the two distributions (with assumed equal covariance), one with mean at $(0,2)$ and the other with mean at $(0,0)$, a straight line through $(0,1)$ would be the reasonable choice. Indeed, we see this in the error rates. The true decision boundary achieved an error rate of $0.0465$, the nonpooled covariance an error rate of $0.0955$, and the pooled covariance an error rate of $0.08$, so the simpler decision boundary actually acheived a better error rate.

Also notable is the fact that the equal-covariance assumption is way off in this case. The two distributions are cigar-shaped, but aligned in perpendicular directions, so with enough samples, the pooled covariance should end up with two very similar eigenvalues. Nevertheless, with so few samples, it's still possible for it to do better than a wildly inaccurate quadratic boundary.
\end{document}