### 2014-04-19 stat246/hw3.tex
s3

\documentclass{article}
\usepackage{geometry}
\usepackage[namelimits,sumlimits]{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage[cm]{fullpage}
\newcommand{\tab}{\hspace*{5em}}
\newcommand{\conj}{\overline}
\newcommand{\dd}{\partial}
\newcommand{\ep}{\epsilon}
\newcommand{\openm}{\begin{pmatrix}}
\newcommand{\closem}{\end{pmatrix}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Null}{null}
\newcommand{\nc}{\newcommand}
\newcommand{\rn}{\mathbb{R}}
\nc{\cn}{\mathbb{C}}
\nc{\ssn}[1]{\subsubsection*{#1}}
\nc{\inner}[2]{\langle #1,#2\rangle}
\nc{\h}[1]{\widehat{#1}}
\nc{\tl}[1]{\widetilde{#1}}
\nc{\norm}[1]{\left\|{#1}\right\|}
\begin{document}

Name: Hall Liu

Date: \today 
\vspace{1.5cm}
\subsection*{1}
\ssn{a}
Let $Y=1,\ldots,M$ with $P(Y=m)=\pi_m$, and let $(X|Y=m)\sim N(\mu_m, \sigma_m^2)$. Then, the joint distribution of $X$ and $Y$ is the product of these,
\[f_{X,Y}(x,y)=\phi(x;\mu_y,\sigma_y^2)\pi_y\]
To get the marginal density of $X$ we sum over $Y$, which gives
\[\sum_{m=1}^M\phi(x;\mu_m,\sigma_m^2)\pi_m\]
where $\phi$ is the normal pdf, so this gives us the density for the mixture model.
\ssn{b}
The likelihood for a sample $(x_i,y_i)$ is 
\[\prod_{i=1}^N\phi(x_i;\mu_{y_i}, \sigma_{y_i}^2)\pi(y_i)=\frac{1}{(2\pi)^{N/2}}\prod_{i=1}^N\frac{1}{\sigma_{y_i}}\exp\left(\frac{(x_i-\mu_{y_i})^2}{2\sigma_{y_i}^2}\right)\pi(y_i)\]
whose logarithm is 
\[-\frac{N}{2}\log(2\pi)\sum_{i=1}^N\left(-\log(\sigma_{y_i})+\log\pi(y_i)+\frac{(x_i-\mu_{y_i})^2}{2\sigma_{y_i}^2}\right)\]
Now, separate it by the value of the $Y_i$, letting $N_m$ be the number of observations with $Y=m$. Doing this and discarding the constant factor in the front, we get
\[\sum_{m=1}^MN_m\left(\log\pi_m-\log\sigma_m\right)+\sum_{i=1}^{N_m}\frac{(x_i-\mu_m)^2}{2\sigma_m^2}\]
Differentating wrt $\pi_m$ and setting equal to $0$, we obtain $\h{\pi_m}=N_m$, which after normalizing so that their sum is $1$, becomes $\h{\pi_m}=\frac{N_m}{N}$. Then, what we obtain is
\[\sum_{m=1}^M\sum_{i=1}^{N_m}\frac{(x_i-\mu_m)^2}{2\sigma_m^2}-\log\sigma_m+\h{\pi_m}\]
Then, we just have a collection of $M$ unrelated normal log-likelihoods to maximize, so the MLE of the means and the variances are just the sample mean and sample variance within the classes.
\ssn{c}

\end{document}